An Integration of Multiple Constraints in ACG

### 2013-07-13 lenls/abstract.tex
LENLS: Changed the title
First rereading of the LENLS abstract

\title{Towards a Wide-Coverage Grammar: \\ Graphical Abstract Categorial Grammars}
\author{Jiří Maršík \and Maxime Amblard}
\maketitle

@@

An ACG is defined in terms of two higher-order signatures, sets of typed
constants on which we can form well-typed lambda terms, and a
translating function from terms of one of the signatures to terms of the
other (the former signature being called the \emph{abstract} signature,
the latter being called the \emph{object} signature and the function
being called a \emph{lexicon}). We are then interested in the terms of
the abstract signature having some distinguished type and their images
through the lexicon (these two sets of terms are called the
\emph{abstract language} and the \emph{object language}, respectively).

When describing language, we are usually interested in systems of more
than one ACG and we use diagrams such as the ones on
Figure~\ref{fig:acg-comp} to explain them.

In Figure~\ref{fig:acg-comp-basic}, we have abstract syntactic terms
from the signature $\Sigma_{Synt}$ yielding the phonological strings and
semantic representations using two different lexicons. In
Figure~\ref{fig:acg-comp-constr}, we extend this pattern by adding a new
abstract signature which narrows the language of syntactic terms and
implements some linguistic constraint. Finally, in
Figure~\ref{fig:acg-comp-sem}, we split the abstract signature
$\Sigma_{Synt}$ of Figure~\ref{fig:acg-comp-basic} into two signatures,
$\Sigma_{Syntax}$ and $\Sigma_{SyntSem}$, in order to model syntactic
and semantic ambiguities in different grammars.

In our grammar, we would like to combine both of the patterns on
Figures~\ref{fig:acg-comp-constr} and \ref{fig:acg-comp-sem} and
furthermore, to use the pattern of Figure~\ref{fig:acg-comp-constr}
multiple times to implement multiple constraints independently. However,
these combinations are not possible in the framework of ACGs without
introducing complexity such as by contaminating the syntax-semantics
interface by the implementation details of the syntactic layer.

@@

We define a \emph{graphical abstract categorial grammar} as a directed
acyclic graph (DAG) whose nodes are labeled with signatures (and
distinguished types) and whose edges are labeled with lexicons, a
mathematical reification of an ACG diagram generalized from
arborescences to DAGs. We then search for an appropriate semantics for
these structures, i.e. how to determine what languages are defined by a
graphical ACG.

We first follow the paradigm in which we associate languages with the
nodes of the graph while the edges tell us how these languages are
defined in terms of each other. We argue that intersection of languages
is the suitable meaning of more than two edges arriving at the same node
based both on the simplicity of the resulting definitions and on our
expectations about the desired semantics. This leads us to the following
definitions of \emph{intrinsic} and \emph{extrinsic} languages
associated with some node $v$ in a graphical ACG $\mathcal{G}$:

@@

The intrinsic language is just the set of terms built on the node's
signature and having the node's distinguished type. The extrinsic
language is established by taking the extrinsic languages of its
predecessors, mapping them through a lexicon and taking their
intersection, or just taking the node's intrinsic language if it has no
predecessors.

We examine the relationship between the languages defined by ACGs and
graphical ACGs (G-ACGs). Intrinsic languages correspond exactly to
abstract languages and therefore the sets of languages definable by both
are equal.

@@

G-ACG extrinsic languages correspond to ACG object languages with
intersection. More formally, we can say that ACG object languages are
just ACG abstract languages closed on transformation by a lexicon and we
show that G-ACG extrinsic languages are just ACG abstract languages
closed on transformation by a lexicon and intersection.

@@

However, extrinsic languages turn out to give counter-intuitive
interpretations to some graphical ACGs and so after examining the
algebraic paradigm which gave us the intrinsic and extrinsic languages,
we consider an alternative paradigm, which we call generative.

In this paradigm, we identify the nodes of the graph with terms and the
edges with statements that one term is mapped into another using a
lexicon. This leads us to the definition of the \emph{pangraphical}
language of a node $u$ in a G-ACG $\mathcal{G}$.

@@

The two paradigms under which we can interpret ACGs are equivalent in
the context of plain ACGs and arborescent diagrams and the two metaphors
(nodes as languages and nodes as terms) can be used interchangeably. We
back this up formally by showing that for every node $u$ in every
arborescent G-ACG $\mathcal{G}$, we have $\mathcal{E}_{\mathcal{G}}(u) =
\mathcal{P}_{\mathcal{G}}(u)$. However, as we start to consider
non-arborescent graphs, we find, interestingly, that the two paradigms
diverge.

The newly defined pangraphical languages solve the problem of extrinsic
languages giving us counter-intuitive interpretations for some specific
G-ACGs. We show by construction that pangraphical languages are as
expressive as extrinsic languages and so the kinds of languages we have
defined here are increasingly more expressive, $\mathcal{I} \subseteq
\mathcal{E} \subseteq \mathcal{P}$. We also show that
$\mathcal{I}_{\mathcal{G}}(u) \supseteq \mathcal{E}_{\mathcal{G}}(u)
\supseteq \mathcal{P}_{\mathcal{G}}(u)$, meaning that the language
definitions are more and more constraining and that the extrinsic
languages went in the right direction as pangraphical languages, but not
as far.

Finally, we construct a G-ACG which integrates the French negation
constraint discussed in Section~\ref{sec:constraints}, the constraints
on extraction introduced in \cite{pogodalla2012controlling} and a
constraint handling gender agreement in a single grammar
specification. The negation constraint is implemented exactly as was
described in Section~\ref{sec:constraints} and the extraction
constraints are taken as is from \cite{pogodalla2012controlling}. The
underlying syntactic signature is defined without any concern for the
additional constraints and the syntax-semantics signature is kept simple
as well.

In the end, this lets us define the syntax in a clean way using the
idiomatic style of categorial grammars (simple atomic types like $N$,
$NP$ and $S$) and then define the constraints themselves the same way
they are defined in papers which try to formalize them individually
(such as the case with \cite{pogodalla2012controlling}). We can freely
combine both the constraint pattern and the syntax-semantics ambiguities
pattern from Section~\ref{sec:acg-comp}.

>>>>

\title{An Integration of Multiple Constraints in ACG}
\author{Jiří Maršík \and Maxime Amblard}
\maketitle

@@

An ACG is defined in terms of two higher-order signatures, sets of typed
constants on which we can form well-typed lambda terms, and a
translating function from terms of one of the signatures to terms of the
other (the former signature being called the \emph{abstract signature},
the latter being called the \emph{object signature} and the function
being called a \emph{lexicon}). We are then interested in the terms of
the abstract signature having some distinguished type and their images
through the lexicon (these two sets of terms are called the
\emph{abstract language} and the \emph{object language}, respectively).

When describing languages, we are usually interested in systems of more
than one ACG and diagrams such as the ones on Figure~\ref{fig:acg-comp}
have been used to explain them.

In Figure~\ref{fig:acg-comp-basic}, abstract syntactic terms from the
signature $\Sigma_{Synt}$ yield phonological strings and semantic
representations using two different lexicons. In
Figure~\ref{fig:acg-comp-constr}, this pattern is extended by adding a
new abstract signature which narrows the language of syntactic terms and
implements some linguistic constraint
\cite{pogodalla2012controlling}. Finally, in
Figure~\ref{fig:acg-comp-sem}, the abstract signature $\Sigma_{Synt}$ of
Figure~\ref{fig:acg-comp-basic} is split into two signatures,
$\Sigma_{Syntax}$ and $\Sigma_{SyntSem}$, in order to model syntactic
and semantic ambiguities in different grammars
\cite{pogodalla2007generalizing}.

In our grammar, we would like to combine both of the patterns on
Figures~\ref{fig:acg-comp-constr} and \ref{fig:acg-comp-sem} and
furthermore, to use the pattern of Figure~\ref{fig:acg-comp-constr}
multiple times to implement multiple constraints independently. However,
these combinations are not possible in the framework of ACGs without
introducing complexity such as contaminating the syntax-semantics
interface by the implementation details of the syntactic layer.

@@

We define a \emph{graphical abstract categorial grammar} as a directed
acyclic graph (DAG) whose nodes are labeled with signatures (and
distinguished types) and whose edges are labeled with lexicons, in other
words just a mathematical reification of an ACG diagram that has also
been generalized from arborescences to DAGs. We then search for an
appropriate semantics for these structures, i.e. how to determine what
languages are defined by a graphical ACG.

We first follow a paradigm in which nodes of the diagrams are
interpreted as languages with the edges telling us how these languages
are defined in terms of each other. A single arrow leading to a language
means that the target language is produced from the source by mapping it
through a lexicon. We then argue that the suitable meaning of two or
more edges arriving at the same node is intersection of languages based
both on the simplicity of the resulting definitions and on our
expectations about the desired semantics. This leads us to the following
definitions of \emph{intrinsic} and \emph{extrinsic} languages
associated with some node $v$ in a graphical ACG $\mathcal{G}$:

@@

The intrinsic language is just the set of terms built on the node's
signature and having the node's distinguished type. The extrinsic
language is established by taking the extrinsic languages of its
predecessors, mapping them through lexicons and taking their
intersection, or just taking the node's intrinsic language if it has no
predecessors.

We then examine the relationship between the languages defined by ACGs
and graphical ACGs (G-ACGs). Intrinsic languages correspond exactly to
abstract languages and therefore the sets of languages definable by both
are equal.

@@

G-ACG extrinsic languages correspond to ACG object languages with
intersection. More formally, it can be said that ACG object languages
are just ACG abstract languages closed on transformation by a lexicon
and we then show that G-ACG extrinsic languages are just ACG abstract
languages closed on transformation by a lexicon and intersection.

@@

However, extrinsic languages turn out to give counter-intuitive
interpretations to some graphical ACGs and we turn to another paradigm.

In the new paradigm, we interpret the nodes of the graph as terms and
the edges with statements that one term is mapped into another using a
lexicon. This leads us to the definition of the \emph{pangraphical}
language of a node $u$ in a G-ACG $\mathcal{G}$.

@@

The two paradigms under which we can interpret ACGs are equivalent in
the context of plain ACGs whose diagrams are arborescent and the two
metaphors (nodes as languages and nodes as terms) can be used
interchangeably. We back this up formally by showing that for every node
$u$ in every arborescent G-ACG $\mathcal{G}$, we have
$\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)$. However,
as we start to consider non-arborescent graphs, we find, interestingly,
that the two paradigms diverge (i.e. $\exists \mathcal{G},
u.\ \mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$).

The newly defined pangraphical languages solve the problem of extrinsic
languages giving us counter-intuitive interpretations for some specific
G-ACGs. We show by construction that pangraphical languages are as
expressive as extrinsic languages and so the kinds of languages we have
defined here are increasingly more expressive, $\mathcal{I} \subseteq
\mathcal{E} \subseteq \mathcal{P}$. We also show that
$\mathcal{I}_{\mathcal{G}}(u) \supseteq \mathcal{E}_{\mathcal{G}}(u)
\supseteq \mathcal{P}_{\mathcal{G}}(u)$, meaning that the language
definitions are more and more constraining and that the extrinsic
languages went in the same direction as pangraphical languages when
narrowing the elements of $\mathcal{I}_{\mathcal{G}}(u)$, but not as
far.

Finally, we construct a G-ACG which integrates the French negation
constraint discussed in Section~\ref{sec:constraints}, the constraints
on extraction introduced in \cite{pogodalla2012controlling} and a
constraint handling agreement in a single grammar specification. The
negation constraint is implemented exactly as was described in
Section~\ref{sec:constraints} and the extraction constraints are taken
``as is'' from \cite{pogodalla2012controlling}. The underlying syntactic
signature is defined without any concern for the additional constraints
and the syntax-semantics signature is kept simple as well.

In the end, this lets us define the syntax in a clean way using the
idiomatic style of categorial grammars (simple atomic types like $N$,
$NP$ and $S$) and then define the constraints themselves the same way as
they are defined in papers which try to formalize them individually
(such as the case with \cite{pogodalla2012controlling}). We can also
freely combine both the constraint pattern and the syntax-semantics
ambiguities pattern from Section~\ref{sec:acg-comp}.
